---
title: "Trial/participant exclusion"
author: '[Guillermo Montero-Melis](https://www.mpi.nl/people/montero-melis-guillermo)'
date: '`r as.character(Sys.Date())`'
output:
  html_document:
    depth: 2
    number_sections: yes
    theme: default
    toc: yes
    code_folding: show
---

Introduction
============

This report applies the second exclusion criterion indicated in our Registered
Report (https://osf.io/26fbh/, p.22) and documents the results:

> 1) We will exclude trials in which the participant starts the oral recall before the beep, that is, if the word onset falls before the end of the 6 second memory phase (see Figure 1).
> 2) In the two interference conditions, we will exclude trials in which participants fail to execute the interference task, which we define as starting the paradiddle later than 3 seconds into the memory phase (that is, if the first tap is registered later than 3 seconds after the offset of the fourth word in the trial).
>
> At the participant level, we will exclude participants for whom either of the above criteria or technical failure (e.g., recording not working) result in excluding more than 30% of the trials across blocks or more than 50% of trials in a single block. All exclusions will take place before the recall data is coded and analysed.

A list of to-be-excluded trials is saved to disk to be read by data processing
scripts before analysis.


Session set up
==============

Libraries:

```{r setup, warning = FALSE, message = FALSE}
library("knitr")
library("tidyverse")
theme_set(theme_bw())  # ggplot theme
knitr::opts_chunk$set(fig.height=4)
```

Source file where functions are defined:

```{r, message = FALSE}
source("myfunctions/paradiddle_fncs.R")
```


An example first
======

Extract tapping information from one participant's log file:

```{r}
log_ex <- extract_tap_info("../data/raw_data/1_sp13_replication_swe_2020_Oct_30_1334.log")
```

Plot taps for this participant. 
For the first practice trials there should be no tapping, so this plot should be
empty:

```{r, fig.height=1}
plot_taps(log_ex, "word_presentation_practice")
```

Block-specific training trials:

```{r, fig.height=2}
plot_taps(log_ex, "word_presentation_training")
```

Real trials:

```{r, fig.height=8}
plot_taps(log_ex, "word_presentation")
```

For exclusion, we *only* care about whether they started the paradiddle before
3 seconds of the memory period had ellapsed. In this case, for instance, trials
2 and 3 from the arm block will have to be excluded.

The following function computes this; we show only the 6 first trials of the
arm block (block 2):

```{r}
compute_valid(log_ex) %>% head(6) %>% kable(digits = 3) 
```



Compute valid/invalid trials for all participants
============================================

Make a list of all valid participant files:

```{r}
# save the name of all log files into vector
path <- "../data/raw_data/"
logfiles <- list.files(path, pattern = ".*\\.log")  # select .log files

# get the participant ID for each logfile
logfiles_id <- gsub("^(\\d+)_.*", "\\1", logfiles)

# load participant sheet where we've marked who needs to be excluded
ppt_data <- read_csv("../data/participant_data_raw.csv") %>%
  select(1:2)
names(ppt_data)[2] <- "valid"
# select valid participants and extract their ID
ppt_data <- ppt_data %>%
  filter(valid %in% c(1, 2)) %>%  # values of 1 and 2 are valid
  mutate(valid_id = gsub("^(\\d+),.*", "\\1", pptID))  # copy ppt ID to column

# vector of valid participant IDs
valid_ids <- unique(ppt_data$valid_id)

# only keep logfiles of valid participants
logfiles_valid <- logfiles[logfiles_id %in% valid_ids]
# exclude also those that contain "excl" (marked by RA to be excluded)
logfiles_valid <- logfiles_valid[! grepl("excl", logfiles_valid)]
```

For each participant, compute how many trials need to be excluded by block:

```{r, message=FALSE}
# create a dataframe of all trials marking if they need to be excluded or not
ppt_valid_trials <- map(logfiles_valid, function(logf) {
  full_path <- paste0(path, logf)
  extract_tap_info(full_path) %>%
    compute_valid()
  }) %>%
  bind_rows()
```

```{r}
# summarize the results by participant
ppt_valid_trials_summary <- compute_valid_summary(ppt_valid_trials)
```


Apply exclusion criteria
========================

Participant-block combinations with more than 50% invalid trials?

```{r}
ppt_valid_trials_summary %>%
  filter(prop_invalid > .5)
```

**No exclusions of entire blocks based on this criterion.**


Participants with more than 30% invalid trials across both critical blocks?

```{r}
ppt_valid_trials_summary %>%
  group_by(ppt_file) %>%
  summarise(prop_invalid = mean(prop_invalid)) %>%
  filter(prop_invalid > .3)
```

**No exclusions of whole participants based on this criterion either.**

**Conclusion**

Only trial-level exclusion need to be done!


Save a file to disk with the trials that need to be excluded
===========================================================

```{r}
# rename columns for compatibility with the data processing script where this
# is going to be used:
excluded_trials <- ppt_valid_trials %>%
  ungroup() %>%
  filter(valid == 0) %>%
  mutate(
    # ID unique is participant number and DateTime as per log file
    ID_unique = gsub(".*ta/(\\d+)_.*_(202\\d.*)\\.log", "\\1_\\2", ppt_file),
    trial_block = trial + 1  # trial_block = 1, ..., 26
  ) %>%
  select(ID_unique, block, trial_block) %>%
  arrange(ID_unique, block, trial_block)
head(excluded_trials)
```

Save to disk:

```{r}
write_csv(excluded_trials, "../data/excluded_trials_bc_of_paradiddle.csv")
```



Appendix: complete lists of invalid trials
=========================

First a list of invalid trials per participant

```{r}
ppt_valid_trials_summary %>%
  group_by(ppt_file) %>%
  summarise(prop_invalid = mean(prop_invalid)) %>%
  arrange(-prop_invalid) %>%
  kable(digits = 3)
```

Next a list of invalid trials per participant-block

```{r}
ppt_valid_trials_summary %>%
  arrange(- prop_invalid) %>%
  kable(digits = 3)
```


